---
title: "Power Analysis"
author: "Tobias Dienlin"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T)
set.seed(1)
options(scipen = 10000)
```

```{r packages, message=FALSE, results='hide'}
library(broom)
library(ggplot2)
library(magrittr)
library(tidyverse)
```

# Background

Here, we run some power analysis for a study on online political participation. 
In the study, people use a social networking site (discord) on which they discuss political matters. 
Participants will communicate in groups of 20 people each (medium scale group communication).
The dependent variable is how much people discuss (measured via _number of words_).
We're interested how different SNS designs affect communication.

The study design is as follows:

 - The website is experimentally manipulated (2 x 2 design). 
 - First, the _persistence_ of the comments is manipulated (permanent vs. ephemeral), as well as _identification_ (identifiable vs. anonymous).
 - To increase generalizability, the groups will discuss one of three _topics_: corona politics, environment, gender.

So these are 2 (identification) x 2 (persistence) experimental factors and a 3-way generalization factor (topic), resulting in a minimum of 12 groups.
Hence, the minimum sample size 12 x 20 = 240 participants.

To calculate power, the question now is how often do we need to _repeat_ this design to achieve sufficient power? Once, twice, thrice, etc? Hence, the factor to change/affect power is _repetition_, ranging from 1 to 10.

# Custom functions

```{r functions}
generate_design <- function(groupsize, 
                            persis, 
                            ident, 
                            topics, 
                            repetition, 
                            ...){
  
  # function generates underlying (empty) datastructure
  
  # count number of groups
  groups <- persis * ident * topics * repetition
  
  # make datastructure
  expand.grid(
    participant = 1:groupsize, 
    persistence = 1:persis - 1, # -1 to make binary
    identification = 1:ident - 1, 
    topic = 1:topics,
    repetition = 1:repetition) %>% 
    as.data.frame() %>% 
    rownames_to_column("id") %>% 
    mutate(
      group = rep(c(1:groups), each = groupsize))
}

sim_d <- function(d_frame, 
                  seed, # make results reproducible
                  effects, # vector of effects we anticipate
                  sd, 
                  groupsize, 
                  ...){
  
  # function to simulate data

  # set.seed(seed)  # uncomment to make results reproducible
  
  # compute how many participants per cell (exp. condition)
  n_cell <- groupsize_n * topics_n * repetition_n
  
  # create the DV. 
  # For now, this will be standardized, bc. of lack of concrete data

  # version 1: if-else # produces weird results
  # d_frame %<>%
  #   mutate(words = 
  #     ifelse(persistence == 0 & identification == 0, rnorm(n_cell, effects["pers0_iden_0_m"], sd),
  #        ifelse(persistence == 0 & identification == 1, rnorm(n_cell, effects["pers0_iden_1_m"], sd),
  #            ifelse(persistence == 1 & identification == 0, rnorm(n_cell, effects["pers1_iden_0_m"], sd),
  #                rnorm(n_cell, effects["pers1_iden_1_m"], sd))))
  #          )
  
  # version 2: filter
  d_frame$words <- 999
  for(i in 1 : repetition_n) {
    d_frame[d_frame$persistence == 0 & d_frame$identification == 0 & d_frame$repetition == i, ]$words <- 
      rnorm(groupsize_n, effects["pers0_iden_0_m"], sd)
    d_frame[d_frame$persistence == 0 & d_frame$identification == 1 & d_frame$repetition == i, ]$words <- 
      rnorm(groupsize_n, effects["pers0_iden_1_m"], sd)
    d_frame[d_frame$persistence == 1 & d_frame$identification == 0 & d_frame$repetition == i, ]$words <- 
      rnorm(groupsize_n, effects["pers1_iden_0_m"], sd)
    d_frame[d_frame$persistence == 1 & d_frame$identification == 1 & d_frame$repetition == i, ]$words <- 
      rnorm(groupsize_n, effects["pers1_iden_1_m"], sd)
  }
    
  return(d_frame)
}

analyze_d <- function(object, ...) {

  # function to analyze data and to extract results
  
  # get means
  means <- group_by(object, persistence, identification) %>% 
    summarize(mean = mean(words), .groups = 'drop')
  
  # get estimates from regression
  fit <- lm(words ~ persistence + identification, object)
  fit_res <- tidy(fit)
  
  # combine result
  results <- data.frame(
    reps = repetition_n,
    per0_ide0_m = filter(means, persistence == 0, identification == 0)$mean,
    per0_ide1_m = filter(means, persistence == 0, identification == 1)$mean,
    per1_ide0_m = filter(means, persistence == 1, identification == 0)$mean,
    per1_ide1_m = filter(means, persistence == 1, identification == 1)$mean,
    persistence_est = fit_res[fit_res$term == "persistence",]$estimate,
    persistence_p = fit_res[fit_res$term == "persistence",]$p.value,
    identification_est = fit_res[fit_res$term == "identification",]$estimate,
    identification_p = fit_res[fit_res$term == "identification",]$p.value,
    n = glance(fit)$nobs
  ) 
  return(results)
}

des_sim_fit <- function(...){
  
  # function to report and extract results
  
  d_frame <- generate_design(...)
  d <- sim_d(d_frame, ...)
  analyze_d(d)
}

est_pow <- function(n_sim = 1000, ...){
  # function to run analyse n_sim times

  tibble(sim = 1:n_sim) %>% 
  mutate(
    effect = map(sim, 
                 des_sim_fit, 
                 groupsize = groupsize_n, 
                 persis = persis_n, 
                 ident = ident_n, 
                 topics = topics_n, 
                 repetition = repetition_n, 
                 effects = effects_est, 
                 sd = sd_est)
    ) %>%
  unnest(effect) %>%
  as.data.frame()
}

```

# Study Design

```{r}
# study design
groupsize_n   <- 20
persis_n      <- 2
ident_n       <- 2 
topics_n      <- 3

# minimum sample size
sample_size <- groupsize_n * persis_n * ident_n * topics_n
```

We define our study design as follows:

- `r groupsize_n` participants per group
- `r persis_n` persistence conditions
- `r ident_n` identification conditions
- `r topics_n` different topics to be discussed
- `r sample_size` minimum sample size

# Define effect size

We then need to define likely effects. Problem is, we don't have good estimates of actual, raw date. To simplify, we assume normal distribution, a mean of zero and a standard deviation of one. 
We can hence think of effects in terms of Cohen's d: .2 = small, .5 = medium, and .8 = large.

|              | persistent | ephemeral |
|--------------|:----------:|:---------:|
| identifiable |    -.40    |   -.20    |
| anonymous    |    -.20    |     0     |

This should lead to a main effect of persistence of d = -.20 and a main effect of identifiability of d = -.20.

```{r}
pers0_iden_0_m <- 0
pers0_iden_1_m <- -.2
pers1_iden_0_m <- -.2
pers1_iden_1_m <- -.4
effects_est <- c(pers0_iden_0_m, pers0_iden_1_m, pers1_iden_0_m, pers1_iden_1_m)
names(effects_est) <- c("pers0_iden_0_m", "pers0_iden_1_m", "pers1_iden_0_m", "pers1_iden_1_m")
sd_est <- 1
```

# Test run

To see if our functions work, let's make a test run with only one repetition.

```{r}
repetition_n <- 1
```


## Set-up

We first create an empty data frame, in which we will then later simulate the data.

```{r results='hide'}
# create design frame
d_frame <- generate_design(
  groupsize  = groupsize_n,
  persis     = persis_n,  
  ident      = ident_n,     
  topics     = topics_n,  
  repetition = repetition_n
  )
d_frame
```

Check if data-frame is alright.

```{r results='hide'}
xtabs(~persistence + identification + topic + repetition, d_frame)
```

Allocation of participants to experimental groups worked just fine.

## Simulate data

Let's create a single data-set and analyze it.

```{r}
d <- sim_d(d_frame, seed = 1, effects_est, sd_est, groupsize_n)
```

## Analyse data

Let's check if means were created alright:

```{r}
means <- d %>% 
  group_by(persistence, identification) %>% 
  summarize(mean = mean(words), .groups = 'drop')
means
```

Sample size small and single study, but general tendency seems to be alright.

Let's also quickly run a regression.

```{r}
lm(words ~ persistence + identification, d) %>% 
  summary()
```

Results look reasonable. Both persistence and identification reduce disclosure.

# Power analysis

Let's next run our actual power analysis.

```{r}
n_sim <- 1000
n_reps <- 10
```

We run a power analysis with `r n_sim` simulations per design. We test 10 designs, that is 1 to `r n_reps` repetitions.

```{r results='hide'}
# create empy data frame
columns <- c("sim", "reps", "per0_ide0_m", "per0_ide1_m", 
             "per1_ide0_m", "per1_ide1_m", "persistence_est", 
             "persistence_p", "identification_est", "identification_p", "n")
sims <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
colnames(sims) = columns

for(i in 1 : n_reps){
  repetition_n  <- i
  sims <- rbind(sims, est_pow())
}
```

## Visualization

Let's inspect the results. First persistence:

```{r}
ggplot(sims) +
  geom_point(aes(sim, persistence_est, color = persistence_p < .05), 
             size = .2, alpha = .5) + 
  scale_color_manual(values = c("darkgrey", "blue")) +
  facet_wrap(facets = "reps")
```

Shows that with more repetitions, effect size move closer to actual population value.

To make sure, let's next check identification -- should provide identical results.

```{r}
ggplot(sims) +
  geom_point(aes(sim, identification_est, color = identification_p < .05), 
             size = .2, alpha = .5) + 
  scale_color_manual(values = c("darkgrey", "blue")) +
  facet_wrap(facets = "reps")
```

Looks good. 

## Cell means & main effects

Next, we compute the average means in the four cells averaged across simulations, plus the two main effects.
This is more of a sanity check to see if our population values can be reproduced.

```{r}
sims %>% 
  group_by(reps) %>% 
  summarise(per0_ide0 = mean(per0_ide0_m),
            per0_ide1 = mean(per0_ide1_m),
            per1_ide0 = mean(per1_ide0_m),
            per1_ide1 = mean(per1_ide1_m),
            persistence = mean(persistence_est), 
            identification = mean(identification_est)
            )
```

Shows that the means resemble those we defined a priori. Same for main effects.

## Power Estimates

Now, let's compute power for each number of replication.

```{r}
sims %>% 
  group_by(reps) %>% 
  summarise(persistence = sum(persistence_p < .05 & persistence_est < 0) / n_sim,
            identification = sum(identification_p < .05 & identification_est < 0) / n_sim,
            n = max(n))
```
If we replicate the study at least 5 times, then we get more than 80% power.

# Next steps

- Run power analysis for Bayes Factors comparing different hypotheses
- Simulate new data with unstandardized results
- Simulate new data with more realistic distribution; i.e., zero-inflated beta/gamma distribution.
- Analyze data using mixed effects model
- Simulate data with hierarchical structure
